---
title: "R Tutorial — variables and functions"
author: "Data Visualization (DS4073)"
date: ""
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Data Types

The variables are assigned with R-Objects and the data type of the R-object becomes the data type of the variable. There are many types of R-objects.

-   Vectors
-   Lists
-   Matrices
-   Arrays
-   Factors
-   Data Frames

The simplest of these objects is the vector object and there are six data types of these atomic vectors. The other R-Objects are built upon the atomic vectors.

```{r}
v <- TRUE 
print(class(v))
v <- 23.5
print(class(v))
v <- 2L
print(class(v))
v <- 2+5i
print(class(v))
v <- "TRUE"
print(class(v))
v <- charToRaw("Hello")
v
print(class(v))
```

# 1.1 Vectors

Atomic vector.

```{r}
# Atomic vector of type double.
print(12.5)
```

Multiple Elements Vector.

1.  Using colon operator with numeric data.
2.  Using sequence (Seq.) operator
3.  Using the c() function

```{r}
# 1. Using colon operator
# Creating a sequence from 5 to 13.
v <- 5:13
print(v)
# Creating a sequence from 6.6 to 12.6.
v <- 6.6:12.6
print(v)
# If the final element specified does not belong to the sequence then it is discarded.]
v <- 3.8:11.4
print(v)

# 2. Using sequence (Seq.) operator
# Create vector with elements from 5 to 9 incrementing by 0.4.
print(seq(5, 9, by = 0.4))

# 3. Using the c() function
# The logical and numeric values are converted to characters.
s <- c('apple','red',5,TRUE)
print(s)
print(class(s[3]))
```

Accessing Vector Elements

$\color{red}{\text{In class exercise}}$

```{r}
# Accessing vector elements using position.
fruits <- c("banana", "apple", "orange", "mango", "lemon")
# Q1: Find out how many items a vector has.
length(fruits)
# Q2: Access the first and third item (banana and orange)
fruits[c(1,3)]
v=seq(1,3,2)
fruits[v]
# Q3: Access all items except for the first and last items.
fruits[c(-1,-length(fruits))]
# Q4: Change "mango" to "pear". Suppose you don't know the index of mango.
fruits[fruits=="mango"] <- "pear"
fruits

```

Vector Element Recycling

```{r}
v1 <- c(3,8,4,5,0,11)
v2 <- c(4,11)
# V2 becomes c(4,11,4,11,4,11)

add.result <- v1+v2
print(add.result)
```

# 1.2 Lists

Lists are the R objects which contain elements of different types like − numbers, strings, vectors and another list inside it.

```{r}
# Create a list containing a vector, a matrix and a list.
list_data <- list(c("Jan","Feb","Mar"), 
                  matrix(c(3,9,5,1,-2,8), 
                         nrow = 2),
                  list("green",12.3))

list_data[[1]][3]
list_data[[2]][1]
list_data[[3]][2]
li=list(1,2,3)
li[[3]]
# Give names to the elements in the list.
#names(list_data) <- c("1st Quarter", "A_Matrix", "A Inner list")
```

$\color{red}{\text{In class exercise}}$

```{r}
# Q1: Access the first element of the list using position.
list_data[1]
# Q2: Access the last element of the list using name.
list_data["A Inner list"]
# Q3: Add element "New element" at the end of the list.
list_data[length(list_data)+1] <- "New element"
#list_data
# Q4: Remove the last element
list_data[length(list_data)] <- NULL
#list_data
#list_data[[3]][[2]]

x <- list(
  a = 1:3,
  b = list("red", c(10,20)),
  c = matrix(1:4, nrow=2)
)
x[[2]][[2]][1]
```

# 1.3 Matrices

The basic syntax for creating a matrix in R is − matrix(data, nrow, ncol, byrow, dimnames)

```{r}
# Elements are arranged sequentially by row.
M <- matrix(c(3:14), nrow = 4, byrow = TRUE)
print(M)

# Define the column and row names.
rownames = c("row1", "row2", "row3", "row4")
colnames = c("col1", "col2", "col3")

P <- matrix(c(3:14), nrow = 4, byrow = TRUE, dimnames = list(rownames, colnames))
print(P)
```

# 1.4 Arrays

Arrays are the R data objects which can store data in more than two dimensions.

```{r}
# Create two vectors of different lengths.
vector1 <- c(5,9,3)
vector2 <- c(10,11,12,13,14)

# Take these vectors as input to the array.
result <- array(c(vector1,vector2),dim = c(2,4,2))
print(result)
print(dim(result))
```

$\color{red}{\text{In class exercise}}$

```{r}
# Q1: create array with dim = c(2,4,3) and print the result
result <- array(c(vector1,vector2),dim = c(2,4,3))
print(result)

# create array with dim = c(2,3,4) and print the result
result <- array(c(vector1,vector2),dim = c(2,3,4))
print(result)

# Q2: Print the second row of the second matrix of the array.
print(result[2,,2])

# Q3: Use apply() function to calculate the sum of the rows across all the matrices.
print(result)
result2 <- apply(result,1,sum)
print(result2)
#Q4: Sort the columns of the matrix
result3 <- apply(result,2,sort)
print(result3)
#print(dim(result3))
```

# 1.5 Factors

Factors are the R-objects which are created using a vector. It stores the vector along with the distinct values of the elements in the vector as labels.

```{r}
# Create a vector
apple_colors <- c('green', 'green', 'yellow', 'red', 'red', 'red', 'green')

# Create a factor object
factor_apple <- factor(apple_colors)

# Print the factor
print(factor_apple)
```

# 1.6 Data Frame

Data frames are tabular data objects. Unlike a matrix in data frame each column can contain different modes of data.

```{r}
# Create the data frame.
emp.data <- data.frame(
   emp_id = c (1:5),
   emp_name = c("Rick","Dan","Michelle","Ryan","Gary"),
   salary = c(623.3,515.2,611.0,729.0,843.25),
   
   start_date = as.Date(c("2012-01-01","2013-09-23","2014-11-15","2014-05-11",
      "2015-03-27")),
   stringsAsFactors = FALSE
)
print(emp.data)

# Extract Specific columns.
result <- data.frame(emp.data$emp_name,emp.data$salary)
print(result)
```

$\color{red}{\text{In class exercise}}$

```{r}
#Q1: Given column vectors, use cbind() to combine multiple vectors column by column to form a new data frame
car_models <- c('Maruti','Hyundai','Tata',
                'Ford','Nissan','Toyota')
car_type <- c('Diesel','Petrol','Petrol',
              'Diesel','Petrol','Diesel')
car_color <- c('Red','Blue','Red',
               'Red','Blue','Red')
year <- c(2001,2011,2013,2012,2021,2021)
# write your answer here:
mat <- cbind(car_models,car_type,car_color,year)
```

# 2. Decision Making

```{r}
# If statement
a <- 33
b <- 200

if (b > a) {
  print("b is greater than a")
} else if (a == b) {
  print("a and b are equal")
} else {
  print("a is greater than b")
}

# For loop
for (x in 1:10) {
  print(x)
}

fruits <- list("apple", "banana", "cherry")
for (x in fruits) {
  if (x == "cherry") {
    break
  }
  print(x)
}
```

# 3. File Handling

In R Programming, handling of files such as reading and writing files can be done by using in-built functions present in R base package.

```{r}
#data=read.table(file.choose(),header=T)
data <- read.table(file = "./country.txt",header=T)
print(is.data.frame(data))
print(ncol(data))
print(nrow(data))
str(data)
print(data)

```

$\color{red}{\text{In class exercise}}$

```{r}
# Q1: print the first 11 rows of data
print(head(data,11))
# Q2: Get the max gdp from data frame.
maxGDP <- max(data$gdp)
print(maxGDP)
# Q3: Get the country detail with max gdp
retval <- subset(data, gdp == max(data$gdp))
print(retval)
# Q4: write the first 20 rows of data to a txt file "country20.txt".
write.table(data[1:20,], file = "country20.txt")
# Q5: write the first 20 rows of data to a csv file "country20.csv".
write.csv(data[1:20,], file = "country20.csv")
# Q6: read country20.csv and check the results
myData = read.csv("country20.csv")
print(myData)
# Q7: What is the mean of GDP for the developing countries which have death rate higher than 10?
attach(data)
mean(gdp[develop==1& deathrate>10])
```

# 4. Functions

-   Built-in Function: Simple examples of in-built functions are seq(), mean(), max(), sum(x) and paste(...) etc.

```{r}
# Sort the elements of the vector.
v <- c(3,8,4,5,0,11, -9, 304)
sort.result <- sort(v)
print(sort.result)
```

-   User-defined Function: They are specific to what a user wants and once created they can be used like the built-in functions.

```{r}
# Create a function to print squares of numbers in sequence.
new.function <- function(a) {
   for(i in 1:a) {
      b <- i^2
      print(b)
   }
}	
# Call the function new.function supplying 6 as an argument.
new.function(6)
```

$\color{red}{\text{In class exercise}}$ Q1: Fibonacci sequence is defined by F(1)=F(2)=1 and F(n)=F(n–1)+F(n–2). Write an R function called Fibonacci which displays the first n Fibonacci numbers and computes their average. For example, if the input n is 20, the output of Fibonacci (20) shall look like: The first 20 Fibonacci numbers are: 1 1 2 3 5 8 13 21 34 55 89 144 233 377 610 987 1597 2584 4181 6765 The average is 885.5

```{r}
Fibonacci<-function(n){
  F=c(1,1)
  if (n>=3)
    {for (i in 3:n)
     F[i]=F[i-1]+F[i-2]
     }
 af=mean(F[1:n])
 cat(paste('The first',n,'Fibonacci numbers are:'),'\n',F[1:n],'\n',paste('the average is',af))
}

Fibonacci(20)
```
