---
title: "Lab10 Dashboard"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: fill
    social: [ "wechat", "menu"]
    source_code: embed
---

```{r setup, include=FALSE}
library(ggplot2)
library(flexdashboard)
```

# PAGE 0 {.sidebar}

Setup: Please install the package `flexdashboard` if you have not installed it.

```{r eval=F}
#pkgs <- c("flexdashboard")
#install.packages(pkgs)
```

Layout setting: `orientation: rows/columns/scroll`

Convert a page or a column into a sidebar on the left side of the dashboard by adding the `{.sidebar}` attribute.\
e.g.\
`# PAGE {.sidebar}`

# PAGE 1

## Row

### chart (1)

```{r}

```

## Row

### chart (2)

```{r}

```

### chart (3)

```{r}

```

# Heading/Tabset

## Row 1

### Headings

`# First-level heading (page)`\
`## Second-level heading (row or column)`\
`### Third-level heading (pane for plot, chart, etc.)`

## Row 2

### Tabset

If you have several components you’d like to display within a row or column then rather than attempting to fit them all on screen at the same time you can lay them out as a tabset.

The tabset example uses class attributes, which are denoted with a `.` (period) prefix.

Many flexdashboard layout options use attributes in the form of a **key-value pair**, such as `data-height=600`, in which the key and value are separated by an `=` (equal sign).

## Row 3 {.tabset data-height="300,"}

`## Row 3 {data-height=300, .tabset}`

### chart (4)

```{r}

```

### chart (5)

```{r}

```

### chart (6)

```{r}

```

# PAGE 3 {data-navmenu="Group Pages"}

### Add a menu

You can group pages into a `menu` within the top navigation bar by adding the attribute `{data-navmenu=}` to the page heading. Be careful - do not include spaces around the equals sign otherwise it will not work!

e.g.\
`# PAGE 3 {data-navmenu="Group Pages"}`\
`# PAGE 4 {data-navmenu="Group Pages"}`

# PAGE 4 {data-navmenu="Group Pages"}

# Tables

## Row 1 {data-height="300"}

### Table

You can include R code chunks that print outputs such as tables. But the output will look best and respond to the window size if you use the `kable()` function from **knitr** to display your tables.

#### Show output in table

```{r}
knitr::kable(iris)
```

## Row 3

### Dynamic table

If you want to show a dynamic table that allows the user to filter, sort, and/or click through `pages` of the data frame, use the package **DT** and it’s function `datatable()`.

```{r}
DT::datatable(iris,
              rownames = FALSE,
              options = list(pageLength = 5, scrollY = FALSE),
              class = "display")
```

# Plot

## Row 1

### Plot

```{r}
p <- ggplot(iris,
            mapping = aes(x = Sepal.Length,
                          y = Sepal.Width,
                          color = Species)) + 
  geom_point()
p
```

### Interactive plot

```{r}
library(plotly)
ggplotly(p)
```

## Row 2

`{.hidden}`: hide the page/row/column

### Plot 1 {.hidden}

```{r}
p <- ggplot(iris,
            mapping = aes(x = Sepal.Length,
                          y = Sepal.Width,
                          color = Species)) + 
  geom_point()
p
```

### Interactive plot {.hidden}

```{r}
library(plotly)
ggplotly(p)
```

# Page Icon {data-icon="fa-list"}

You can add an icon to the navigation menu for a page using the `data-icon` attribute, e.g. `{data-icon="fa-list"}`

# Storyboard {.storyboard}

Storyboard page

You can layout one or more pages of a dashboard as storyboards and then use traditional row and column layout on the other pages. To do this, you exclude the `storyboard: true` option and instead add the `{.storyboard}` class to pages you want laid out as storyboards.

### Frame 1 - Commentary

You may wish to add commentary alongside the frames in your storyboard. To do this you add a horizontal rule separator (\*\*\*) after the main body of frame content. Content after the separator will appear in a commentary sidebar to the right. For example:

```{r}
```

------------------------------------------------------------------------

Some commentary about Frame 1.

### Frame 2

```{r}
```

------------------------------------------------------------------------

Some commentary about Frame 2.

### Frame 3

```{r}
```

# Theme

### Theme

A variety of themes are available to modify the base appearance of flexdashboard. Available themes include:

-   default\
-   cosmo\
-   bootstrap\
-   cerulean\
-   journal\
-   flatly\
-   readable\
-   spacelab\
-   united\
-   lumen\
-   paper\
-   sandstone\
-   simplex\
-   yeti

# Resource

<https://pkgs.rstudio.com/flexdashboard/articles/using.html#icon-sets>
