---
title: "Lab3-EDA 4-plot"
author: "SID_Name"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1 4-Plot

## 1.1 Run sequence plot ($Y_i$ versus $i$)

```{r}
# fixed location and fixed variation
n <- 300
y <- rnorm(n,mean=0,sd=5)
plot(1:n,y,type="l",
     main = "Run sequence plot",
     xlab = "i",
     ylab = "Y_i")
```

```{r}
# shifts in location
n <- 200
m <- 100
sdata <- c(rnorm(n,mean=0,sd=2), rnorm(m,mean=-15,sd=2)+
             c(seq(10,0,length.out=50),seq(0,5,length.out=50)))
plot(1:(n+m),sdata,type="l",
     main = "Run sequence plot",
     xlab = "i",
     ylab = "Y_i")
```

## 1.2 Lag plot ($Lag_k(Y)$ vs $Y$)

```{r}
# Random data
n <- 200
# Generates random numbers uniformly distributed between a and b.
y <- runif(n,0,10)
# Adjusts the size of the plot symbols (points). 0.5 makes them smaller than the default.
plot(1:n, y, type = "o", cex=0.5, col="blue")
# Specifies the plotting character (symbol) for the points. 16 corresponds to a filled circle.
points(1:n, y, col="red", cex=0.5, pch=16)
```

```{r}
plot(y[-1], y[-length(y)], type = "p", cex = 1, col = "red",
     pch = 16, main = "Lag 1 Plot of Series",
     xlab = "Lag 1 series",
     ylab = "Series",
     xlim = c(0, 10),
     ylim = c(0, 10))
# y[-1] and y[-length(y)] form a lag 1 plot, where each point represents the pair (y_t, y_{t-1})
```

```{r}
# Data with weak autocorrelation
n <- 200
y <-stats::filter(rnorm(n), filter=rep(1,times=5), circular=TRUE)
plot(1:n, y, type = "o", cex=0.5, col="blue")
points(1:n, y, col="red", cex=0.5, pch=16)
plot(y[-1],y[-length(y)], type="p", cex=1, col="red", pch=16)
```

```{r}
# Data with high autocorrelation
n <- 200
y <- 1:n + rnorm(n,0,2)

plot(1:n, y, type = "o", cex=0.5, col="blue")
points(1:n, y, col="red", cex=0.5, pch=16)
plot(y[-1],y[-length(y)], type="p", cex=1, col="red", pch=16)
```

```{r}
# Sinusoidal data
n <- 200
y <- 2*sin(1:n)+c(rnorm(n-3,0,0.2),rnorm(3,2,3))

plot(1:n, y, type = "o", cex=0.5, col="blue")
points(1:n, y, col="red", cex=0.5, pch=16)
plot(y[-1],y[-length(y)], type="p", cex=1, col="red", pch=16)
```

## 1.3 Histogram

```{r}
# Normal Distribution, Beta Distribution, Chi-Squared Distribution, Gamma Distribution, t-Distribution, Weibull Distribution, Geometric Distribution, Poisson Distribution, Cauchy Distribution
par(mfrow=c(3,3))
n <- 2000
x1 <- rnorm(n,0,1)
x2 <- rbeta(n, 10,1)
x3 <- rchisq(n,3)
x4 <- rgamma(n,5)
x5 <- rt(n,5)
x6 <- rweibull(n, 1)
x7 <- rgeom(n, 0.3)
x8 <- rpois(n, 5)
x9 <- rcauchy(n,1000)
hist(x1, breaks = 50)
hist(x2, breaks = 50)
hist(x3, breaks = 50)
hist(x4, breaks = 50)
hist(x5, breaks = 50)
hist(x6, breaks = 50)
hist(x7, breaks = 50)
hist(x8, breaks = 50)
hist(x9, breaks = 50)

### Task: Change the title of each histogram to corresponding distribution, for example "Normal Distribution"
### write your answer here

```

## 1.3 Normal probability plot (Theoretical normal value vs. ordered Y)

```{r}
# normal data
n <- 200
x1 <- rnorm(n,0,0.1)
qqnorm(x1, main="Normal data")
qqline(x1)
```

A Q-Q plot is a graphical tool to compare the quantiles of your data against the quantiles of a theoretical normal distribution. If your data is normally distributed, the points in the Q-Q plot will approximately lie on a straight line.

How to get theoretical quantiles?

1.  Sort sample data. The number of samples is n.
2.  Calculate the ranking percentile of sample data points.\
    Suppose a sample point is ith point after sorting, then p = i/(n+1).
3.  Calculate the theoretical quantile.\
    The Z value corresponding to the given percentile p under the normal distribution.

```{r}
# non-normal data
n <- 200
x2 <- rweibull(n, 1)
qqnorm(x2, main="Weibull data")
qqline(x2)
```

# 2 In-Class Exercise

Please draw the 4-plot for the data Data.txt (download from iSpace).

Hint: use `par(mfrow=c(2,2))` to separate the window into 2\*2 parts.

```{r}
### write your answer here
```

```{r,echo=FALSE}
data <- as.matrix(read.table("data.txt"))
num_rows <- nrow(data)
par(mfrow=c(2,2))
plot(1:num_rows,data,type="l",xlab="Run sequence plot",ylab="")
plot(data[-1], data[-num_rows], type = "p", cex=0.5, pch=16, xlab = "Lag 1 plot",ylab="")
hist(data, breaks=20, xlab = "Histogram", ylab="", xlim=c(-2000,2000))
qqnorm(data,xlab="Normal probablility plot")
qqline(data)
```
