---
title: 'R Tutorial - Basic'
output: 
  html_document:
    css: style.css
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Prerequisites:**

$\bullet$ Download R

$\bullet$ Download R studio

$\bullet$ Geometrics: How our data being displayed using point, line,
histogram, bar, boxplot

If you have no experience with the R language, here are some resources
for getting started:

$\bullet$ A (very) short introduction to R:
<https://cran.r-project.org/doc/contrib/Torfs+Brauer-Short-R-Intro.pdf>

$\bullet$ Getting up to speed with R:
<https://rkabacoff.github.io/datavis/Rintro.pdf>

# 1. R Overview

R is a programming language and software environment for statistical
analysis, graphics representation and reporting. The core of R is an
interpreted computer language which allows branching and looping as well
as modular programming using functions.

**Rstudio** is an Integrated Development Environment (IDE) for R. There
are two free versions *Rstudio Desktop* and *Rstudio Server*. You can
download it from <https://www.rstudio.com/products/rstudio/download/>.

## 1.1 Feature of R

The following are the important features of R:

$\bullet$ R is a well-developed, simple and effective programming
language which includes conditionals, loops, user defined recursive
functions and input and output facilities.

$\bullet$ R has an effective data handling and storage facility,

$\bullet$ R provides a suite of operators for calculations on arrays,
lists, vectors and matrices.

$\bullet$ R provides a large, coherent and integrated collection of
tools for data analysis.

$\bullet$ R provides graphical facilities for data analysis and display
either directly at the computer or printing at the papers.

# Now we start writing code!

$\bullet$ Create R script and R Markdown in RStudio

File -\> New File -\> R Script, the short cut key is Ctrl+Shift+N.

File -\> New File -\> R Markdown.

Download provided *.rmd* and *style.css* files from ispace, put them in
the same folder, and load *.rmd* in RStudio.

$\bullet$ RMarkdown tips:

1.  How to generate html from .Rmd? Ans: Knit to HTML.

2.  You can quickly insert chunks like these into your file with

-   the keyboard shortcut Ctrl + Alt + I (OS X: Cmd + Option + I)
-   the Add Chunk command in the editor toolbar
-   or by typing the chunk delimiters \`\`\`{r} and \`\`\`.

# 2. Basic Syntax

As a convention, we will start learning R programming by writing a
“Hello, World!” program.

```{r}
# My first program in R Programming
myString <- "Hello, World!"
print (myString)
```

Save the above code in a file test.R and execute it at Linux command
prompt as given below. Even if you are using Windows or other system,
syntax will remain same.

Specifically, we first make sure the current path in the **Terminal**
window is the path of test.R. Then type in the following command prompt
in the **Terminal** window.

```{r,eval=FALSE}
$ Rscript test.R 
```

## 2.1 Comments

Comments are like helping text in your R program and they are ignored by
the interpreter while executing your actual program. Single comment is
written using \# in the beginning of the statement as follows −

```{r}
# Comments: Add two numbers
print(7 + pi)
```

## 2.2 Navigating directories

$\bullet$ getwd(): Returns the current working directory.

$\bullet$ setwd(): Set the working directory.

$\bullet$ dir(): Return the list of the directory.

$\bullet$ sessionInfo(): Return the session of the windows.

$\bullet$ date(): Return the current date.

```{r, eval=FALSE}
# Show directory, session information, date
getwd()
dir()
sessionInfo()
date()
```

## 2.3 Installing packages

```{r, eval=FALSE}
# install packages
install.packages("rmarkdown")
install.packages("knitr")
install.packages("lattice")
install.packages("ggplot2")
# load the required package
library(lattice)
# Use help() function for on-line help
help(print)
```

## 2.4 Clearing the console and environment

We clear console in R and RStudio, In some cases when you run the codes
using “source” and “source with echo” your console will become messy.
And it is needed to clear the console. So let’s now look at how to clear
the console. The console can be cleared using the shortcut key ctrl + L.

```{r, eval=FALSE}
# Define a variable a=1
a <- 1
# Remove a variable a
rm(a)
# Remove all the variables in the current environment
rm(list=ls())
```

## 2.5 Assignment commands

In R, the assignment can be denoted in three ways:

1.  = (Simple Assignment)
2.  \<- (Leftward Assignment)
3.  -\> (Rightward Assignment)

This is an example:

```{r}
# Assignment values to three variables in three ways
var1 = "Simple Assignment"
var2 <- "Leftward Assignment"
"Rightward Assignment" -> var3

# print the three variables
print(var1)
print(var2)
print(var3)

```

# 3. Data Types

The variables are assigned with R-Objects and the data type of the
R-object becomes the data type of the variable. There are many types of
R-objects.

-   Vectors
-   Lists
-   Matrices
-   Arrays
-   Factors
-   Data Frames

The simplest of these objects is the vector object and there are six
data types of these atomic vectors. The other R-Objects are built upon
the atomic vectors.

```{r}
v <- TRUE 
print(class(v))
v <- 23.5
print(class(v))
v <- 2L
print(class(v))
v <- 2+5i
print(class(v))
v <- "TRUE"
print(class(v))
v <- charToRaw("Hello")
print(class(v))
```

## 3.1 Vectors

Atomic vector.

```{r}
# Atomic vector of type double.
print(12.5)
```

Multiple Elements Vector.

1.  Using colon operator with numeric data.
2.  Using sequence (Seq.) operator
3.  Using the c() function

```{r}
# 1. Using colon operator
# Creating a sequence from 5 to 13.
v <- 5:13
print(v)
# Creating a sequence from 6.6 to 12.6.
v <- 6.6:12.6
print(v)
# If the final element specified does not belong to the sequence then it is discarded.]
v <- 3.8:11.4
print(v)

# 2. Using sequence (Seq.) operator
# Create vector with elements from 5 to 9 incrementing by 0.4.
print(seq(5, 9, by = 0.4))

# 3. Using the c() function
# The logical and numeric values are converted to characters.
s <- c('apple','red',5,TRUE)
print(s)
print(class(s[3]))



```

Accessing Vector Elements

$\color{red}{\text{In class exercise}}$

```{r}
# Accessing vector elements using position.
fruits <- c("banana", "apple", "orange", "mango", "lemon")
# Q1: Find out how many items a vector has.
length(fruits)
# Q2: Access the first and third item (banana and orange)
fruits[c(1,3)]
# Q3: Access all items except for the first and last items.
fruits[c(-1,-length(fruits))]
# Q4: Change "mango" to "pear". Suppose you don't know the index of mango.
fruits[fruits=="mango"] <- "pear"
fruits

print(typeof(fruits))
class(fruits)
```

Vector Element Recycling

```{r}
v1 <- c(3,8,4,5,0,11)
v2 <- c(4,11)
# V2 becomes c(4,11,4,11,4,11)

add.result <- v1+v2
print(add.result)
```

We will continue Sec. 3 next week.
