---
title: "Assignment 2"
author: "Junya YANG 2230026188"
date: "`r Sys.Date()`"

output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Description

Please complete the codes such that they can get the same outputs. Submit the **.html** and **.rmd** documents to iSpace.

## **Q1 (15 marks) Use the dataset CPS85 from the package mosaicData. Please count the number of females and males in each sector.**

### **Hint**: You can copy the command `View(CPS85)` in the console to view the tabular data of CPS85.

```{r}
library(dplyr)
library(ggplot2)
data(CPS85, package = "mosaicData")
#View(CPS85)

new_data = CPS85 %>%
  group_by(sex, sector) %>%
  summarise(n = n()) %>%
  arrange(sector, sex)

new_data
```

Construct a stacked bar chart to show the number of females and males in each sectors.

**Hint**: `geom_text()` may help you show the numbers in the bars. The two color values `#C73C68` and `B6C468` can help.

```{r}

ggplot(new_data, aes(x = reorder(sector, -n, sum), y = n, fill = sex)) +
  geom_col(position = "stack") +
  geom_text(aes(label = n),
            position = position_stack(vjust = 0.5),
            color = "white",
            size = 4) +
  scale_fill_manual(values = c(
    "F" = "#C73C68",
    "M" = "#B6C468"   # 注意这里要有 #
  ))  +
  labs(x =NULL,
        y = "Number of persons",
       fill = "Sex") +
  theme(
  panel.grid = element_blank(),
  panel.background = element_blank(),
  axis.line = element_line()
)+
  coord_flip()


```

## **Q2 (15 marks) Based on the dataset `SaratogaHouses` from package `mosaicData`, please draw a scatterplot for house ages and prices with quadratic line of best fit.**

**Hint**: `geom_smooth()` may help you to construct the quadratic regression line. You can specify the points with `size = 1`, `alpha = 0.4`, the fitting line with `size = 0.5`, `color = "tomato"`.

```{r}
library(ggplot2)
library(mosaicData)
library(scales)

ggplot(SaratogaHouses, aes(x = age, y = price)) +
  geom_point(size = 1, alpha = 0.4,color="blue") +
  geom_smooth(
    method  = "lm",
    formula = y ~ poly(x, 2),  
    color   = "tomato",
    size    = 0.5,
    se      = TRUE
  ) +
  scale_y_continuous(labels = label_dollar())+
  labs(
    title = "Age vs. Price",
    subtitle = "Houses in Saratoga",
    x = "Age of house",
    y = "Price"
  )+  
  theme(
  panel.background = element_rect(fill = "white"),  
  panel.grid.major = element_line(color = "grey95"),  
  panel.grid.minor = element_line(color = "grey95")
)

```

## **Q3 (15 marks) Based on the dataset `SaratogaHouses` from package `mosaicData`, please draw the boxplot for the house price by number of rooms.**

**Hint**: Transparency of boxplot is specified as `alpha = 0.5`.

```{r}
library(ggplot2)
library(mosaicData)
library(scales)

ggplot(SaratogaHouses, aes(x = factor(rooms), y = price)) +
  geom_boxplot(
    notch = TRUE,
    fill  = "cornflowerblue",
    alpha = 0.5        
  ) +
  scale_y_continuous(labels = label_dollar()) + 
  labs(
    title = "House Price by Number of Rooms",
    x = "Number of Rooms",
    y = "Price"
  ) +
  theme(
    panel.background = element_rect(fill = "white"),
    panel.grid.major = element_line(color = "grey95"),
    panel.grid.minor = element_line(color = "grey98")
  )

```

## Q4 (15 marks) Please plot the Cleveland dot chart for the percentage of receipts in GDP by country. The datasset newgdp from the package CGPfunctions records current receipts of fifteen national governments as a percentage of gross domestic product. You can read more detail by running the code ??newgdp in the console.

1.  Selecting all the data in year 1979 and mutate the GDP into percentage.
2.  Plot the Cleveland dot chart for GDP percentage by country.

Hint: Please install the package `CGPfunctions` if you haven’t installed. (No need to write the install code in the R code chunk below.) The range of x is (0.2, 0.6).

```{r}
library(dplyr)
library(ggplot2)
library(CGPfunctions)
library(scales)

data(newgdp, package = "CGPfunctions")

gdp_1979 <- newgdp %>%
  filter(Year == "Year1979") %>%
  mutate(GDP_pct = GDP / 100) %>%
  arrange(GDP_pct) %>%                        
  mutate(Country = factor(Country, levels = Country))  

ggplot(gdp_1979,
       aes(x = GDP_pct,
           y = reorder(Country, GDP_pct))) + 
  geom_segment(aes(x = 0.2, xend = GDP_pct,
                   y = Country, yend = Country),
               color = "grey85") +
  geom_point(size = 2,color="pink") +
  scale_x_continuous(
    limits = c(0.2, 0.6),
    labels = label_percent(accuracy = 1)
  ) +
  labs(
    title = "Percentage of Receipts in GDP by Country",
    x = NULL,
    y = NULL
  ) +
  theme_minimal() +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor  = element_blank()
  )

```

## **Q5 (40 marks) A tiny project**

The data file **Real-Estate.txt** contains information on the homes sold in the Denver area during the year 2003. The variables in this data file are as follows:

```{r}
library(readr)
estate_data=readr::read_table("Real-Estate.txt")

View(estate_data)
```

1.  (10 marks) Please analyze the data and abstract at least 2 tasks.

```{r}
library(dplyr)
library(readr)

# Load and lightly clean the data
estate_raw <- read_table("Real-Estate.txt", col_types = cols())
estate <- estate_raw %>%
  mutate(
    Pool     = factor(Pool, labels = c("No pool", "Pool")),
    Garage   = factor(Garage, labels = c("No garage", "Garage")),
    Township = factor(Township)
  )

# Quick structure + summary to understand ranges
glimpse(estate)
summary(estate)

# Task 1: How strongly do home size and bedroom count explain sale price?
task1_overview <- estate %>%
  group_by(Bedrooms) %>%
  summarise(mean_price = mean(Price), mean_size = mean(Size), .groups = "drop")

# Task 2: Do amenities (pool/garage) offset the distance-to-downtown penalty on price?
task2_overview <- estate %>%
  group_by(Pool, Garage) %>%
  summarise(mean_price = mean(Price), med_distance = median(Distance), n = n(), .groups = "drop") %>%
  arrange(desc(mean_price))

task1_overview
task2_overview
```

2.  (15 marks) Design your visualization according to the design principles and give your justification. Please write your design thoughts clearly.

```{r}
cat(
  "- Goal 1: Show how size/bedrooms drive price. Use a scatterplot of Size vs Price; color by Bedrooms to keep a small legend and add one global linear fit to convey trend.\n",
  "- Goal 2: Show whether amenities overcome distance penalty. Plot Price vs Distance; color by pool/garage combinations; smooth with loess to reveal non-linear decay and any uplift from amenities.\n",
  "Design choices: minimalist grid for readability, dollar formatting on price, subdued palette with one accent per category, semi-transparent points to manage overlap, short titles that state the takeaway."
)
```

3.  (15 marks) Implement your visualization by R. Please write the annotations for each step.

```{r}
library(ggplot2)
library(scales)

# Scatter: size vs price, colored by bedrooms, with overall linear fit
p_size_price <- ggplot(estate, aes(x = Size, y = Price, color = factor(Bedrooms))) +
  geom_point(alpha = 0.6, size = 2) +
  geom_smooth(method = "lm", se = FALSE, color = "grey30", size = 0.6) +
  scale_color_brewer(palette = "Set2", name = "Bedrooms") +
  scale_y_continuous(labels = label_dollar(suffix = "k", prefix = "$")) +
  labs(
    title = "Larger homes command higher prices",
    subtitle = "Denver sales, 2003; linear fit across all homes",
    x = "Size (sq ft)",
    y = "Sale price (USD thousands)"
  ) +
  theme_minimal()

# Prep amenity combo label for second plot
estate <- estate %>%
  mutate(amenity = interaction(Pool, Garage, sep = " + "))

# Scatter: distance vs price, colored by amenity combo, with loess smoothing
p_distance_amenity <- ggplot(estate, aes(x = Distance, y = Price, color = amenity)) +
  geom_point(alpha = 0.6, size = 2) +
  geom_smooth(method = "loess", se = FALSE, size = 0.6) +
  scale_color_manual(
    values = c("#6C8EBF", "#F28E2B", "#8FD694", "#C67ACE"),
    name = "Amenities"
  ) +
  scale_y_continuous(labels = label_dollar(suffix = "k", prefix = "$")) +
  labs(
    title = "Pools/garages soften the distance penalty",
    subtitle = "Sale price vs. miles from city center",
    x = "Distance to downtown (miles)",
    y = "Sale price (USD thousands)"
  ) +
  theme_minimal()

p_size_price
p_distance_amenity
```
