---
title: "Beijing Air Quality Comprehensive Dashboard"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: fill
    social: [""]
    source_code: embed
---

```{r setup, include=FALSE}
library(ggplot2)
library(flexdashboard)
library(tidyverse)
library(lubridate)
library(highcharter)
library(plotly)
library(leaflet)
library(geosphere)
library(GGally)
library(scales)
library(DT)
```

# Introduction {.sidebar}

## Setup Instructions

Please install required packages if you haven't:

```{r eval=FALSE}
pkgs <- c("flexdashboard", "tidyverse", "lubridate", "highcharter", 
          "plotly", "leaflet", "geosphere", "GGally", "scales", "DT")
install.packages(pkgs)
```

## About This Dashboard

This comprehensive dashboard integrates three analytical perspectives:

1.  **Time Series Analysis**: Trends, seasonality, and temporal patterns
2.  **Spatial Analysis**: Geographic distribution and regional variations
3.  **Wind Mechanisms**: Wind direction impacts on air quality

**Data**: Beijing air quality monitoring data (2013-2017)

**Navigate**: Use the top menu to explore different analytical pages

# Overview & Trends

```{r data_prep}
# 读取数据
aq <- read_csv("beijing_air_processed.csv", show_col_types = FALSE)

# 数据预处理
aq <- aq %>%
  mutate(
    date = as_date(datetime),
    ym = floor_date(datetime, "month"),
    week = isoweek(datetime),
    season = factor(season, levels = c("Spring", "Summer", "Autumn", "Winter"), ordered = TRUE),
    station = as.factor(station),
    region = as.factor(region),
    season_year = as.integer(season_year),
    wd = factor(wd, levels = c("N","NNE","NE","ENE","E","ESE","SE","SSE","S","SSW","SW","WSW","W","WNW","NW","NNW","CV"))
  )

# 基础统计
daily_pm25 <- aq %>%
  group_by(date) %>%
  summarise(pm25_mean = mean(pm2_5, na.rm = TRUE))

pm_stats <- daily_pm25 %>%
  summarise(
    total_days = n(),
    exceed_days = sum(pm25_mean > 75, na.rm = TRUE),
    exceed_rate = exceed_days / total_days,
    pm25_mean = mean(pm25_mean, na.rm = TRUE),
    pm25_min = min(pm25_mean, na.rm = TRUE),
    pm25_max = max(pm25_mean, na.rm = TRUE)
  )
```

## Row {data-height="200"}

### Min PM2.5

```{r}
valueBox(round(pm_stats$pm25_min, 2), color = "aqua", icon = "fa-cloud")
```

### Max PM2.5

```{r}
valueBox(round(pm_stats$pm25_max, 2), color = "warning", icon = "fa-cloud")
```

### Average PM2.5

```{r}
valueBox(round(pm_stats$pm25_mean, 2), color = "yellow", icon = "fa-bar-chart")
```

### Total Days

```{r}
valueBox(pm_stats$total_days, icon = "fa-line-chart", color = "aqua")
```

### Exceed Days (≥75)

```{r}
valueBox(pm_stats$exceed_days, icon = "fa-times", color = "warning")
```

### Exceed Rate (≥75)

```{r}
valueBox(scales::percent(pm_stats$exceed_rate, accuracy = 0.1), icon = "fa-times", color = "warning")
```

## Row

### Annual PM2.5 Trend

```{r}
annual_pm25 <- aq %>%
  group_by(year) %>%
  summarise(pm25_mean = mean(pm2_5, na.rm = TRUE))

highchart() %>%
  hc_title(text = "Annual PM2.5 Trend") %>%
  hc_xAxis(categories = annual_pm25$year) %>%
  hc_yAxis(title = list(text = "Average PM2.5 (µg/m³)")) %>%
  hc_add_series(name = "PM2.5", data = round(annual_pm25$pm25_mean, 2), type = "line")
```

### Seasonal PM2.5 Variation

```{r}
season_pm25 <- aq %>%
  group_by(season_year, season) %>%
  summarise(pm25_mean = mean(pm2_5, na.rm = TRUE), .groups = "drop")

highchart() %>%
  hc_title(text = "Seasonal PM2.5 Trend") %>%
  hc_xAxis(categories = paste(season_pm25$season_year, season_pm25$season, sep = "-")) %>%
  hc_yAxis(title = list(text = "Average PM2.5 (µg/m³)")) %>%
  hc_add_series(name = "PM2.5", data = round(season_pm25$pm25_mean, 2), type = "line")
```

### Weekly Cycle

```{r}
aq_weekday <- aq %>%
  mutate(weekday = wday(date, label = TRUE, week_start = 1, locale = "C")) %>%
  group_by(weekday) %>%
  summarise(mean_pm25 = mean(pm2_5, na.rm = TRUE))

highchart() %>%
  hc_title(text = "Weekly PM2.5 Cycle") %>%
  hc_xAxis(categories = aq_weekday$weekday, title = list(text = "Weekday")) %>%
  hc_yAxis(title = list(text = "Average PM2.5 (µg/m³)")) %>%
  hc_add_series(type = "line", data = aq_weekday$mean_pm25, name = "PM2.5") %>%
  hc_tooltip(pointFormat = "Day: {point.category}<br>PM2.5: {point.y} µg/m³")
```

## Row

### Calendar Heatmap (2015)

```{r}
calendar_data <- daily_pm25 %>%
  mutate(
    year = year(date),
    month = month(date, label = TRUE, abbr = TRUE),
    day = day(date),
    wday = wday(date, week_start = 1),
    week = isoweek(date),
    pm25_mean = round(pm25_mean, 1)
  )

target_year <- 2015
df <- calendar_data %>% filter(year == target_year)
min_week <- min(df$week)
df <- df %>% mutate(week = week - min_week + 1)

data_list <- list_parse2(df %>% transmute(x = week, y = wday, value = pm25_mean, name = as.character(date)))

highchart() %>%
  hc_title(text = paste("Calendar Heatmap of PM2.5 -", target_year)) %>%
  hc_xAxis(title = list(text = "Week of Year"), tickInterval = 1) %>%
  hc_yAxis(title = list(text = "Weekday"), categories = c("Mon","Tue","Wed","Thu","Fri","Sat","Sun"), reversed = TRUE) %>%
  hc_add_series(type = "heatmap", data = data_list, name = "PM2.5", borderWidth = 0.5, dataLabels = list(enabled = FALSE)) %>%
  hc_colorAxis(min = min(df$pm25_mean), max = max(df$pm25_mean), stops = color_stops(n = 7, colors = c("#f7fbff","#deebf7","#9ecae1","#6baed6","#3182bd","#08519c"))) %>%
  hc_tooltip(formatter = JS("function() { return '<b>' + this.point.name + '</b><br>PM2.5: ' + this.point.value; }"))
```

### Diurnal Cycle (24-hour)

```{r}
hour_pm25 <- aq %>%
  group_by(hour) %>%
  summarise(mean_pm25 = mean(pm2_5, na.rm = TRUE))

highchart() %>%
  hc_title(text = "Diurnal Cycle of PM2.5") %>%
  hc_xAxis(categories = 0:23, title = list(text = "Hour of Day")) %>%
  hc_yAxis(title = list(text = "PM2.5")) %>%
  hc_add_series(data = hour_pm25$mean_pm25, name = "PM2.5", type = "line") %>%
  hc_plotOptions(series = list(marker = list(enabled = TRUE))) %>%
  hc_tooltip(pointFormat = "Hour: {point.category}<br>PM2.5: {point.y}")
```

## Row

### Yearly Exceed Rate

```{r}
daily_pm25_exceed <- daily_pm25 %>%
  mutate(year = year(date), exceed = pm25_mean > 75)

yearly_exceed <- daily_pm25_exceed %>%
  group_by(year) %>%
  summarise(
    total_days = n(),
    exceed_days = sum(exceed, na.rm = TRUE),
    exceed_rate = exceed_days / total_days
  )

exceed_rate_pct <- round(yearly_exceed$exceed_rate * 100, 1)

highchart() %>%
  hc_title(text = "Annual PM2.5 Exceedance Rate") %>%
  hc_xAxis(categories = yearly_exceed$year, title = list(text = "Year")) %>%
  hc_yAxis(title = list(text = "Exceedance Rate (%)"), labels = list(format = "{value}%")) %>%
  hc_colorAxis(min = min(exceed_rate_pct), max = max(exceed_rate_pct), stops = color_stops(colors = c("lightyellow", "orange", "red", "darkred"))) %>%
  hc_add_series(type = "column", name = "Exceedance Rate", data = exceed_rate_pct, colorByPoint = TRUE) %>%
  hc_tooltip(pointFormat = "Exceedance Rate: {point.y}%") %>%
  hc_plotOptions(column = list(dataLabels = list(enabled = TRUE, format = "{y}%")))
```

### Monthly Exceed Rate

```{r}
daily_pm25_month <- daily_pm25 %>%
  mutate(month = month(date), exceed = pm25_mean > 75)

monthly_exceed <- daily_pm25_month %>%
  group_by(month) %>%
  summarise(
    total_days = n(),
    exceed_days = sum(exceed, na.rm = TRUE),
    exceed_rate = exceed_days / total_days
  )

exceed_rate_pct_m <- round(monthly_exceed$exceed_rate * 100, 1)

highchart() %>%
  hc_title(text = "Monthly PM2.5 Exceedance Rate") %>%
  hc_xAxis(categories = month.abb, title = list(text = "Month")) %>%
  hc_yAxis(title = list(text = "Exceedance Rate"), labels = list(format = "{value}%")) %>%
  hc_colorAxis(min = min(exceed_rate_pct_m), max = max(exceed_rate_pct_m), stops = color_stops(colors = c("lightyellow", "orange", "red", "darkred"))) %>%
  hc_add_series(type = "column", name = "Exceedance Rate", data = exceed_rate_pct_m, colorByPoint = TRUE) %>%
  hc_tooltip(pointFormat = "Exceedance Rate: {point.y}%") %>%
  hc_plotOptions(column = list(dataLabels = list(enabled = TRUE, format = "{y}%")))
```

# Spatial Distribution

```{r spatial_prep}
# 读取站点信息
stations <- read_csv("stations.csv", show_col_types = FALSE)

# 合并经纬度
df_spatial <- aq %>%
  left_join(stations, by = "station") %>%
  mutate(region = region.y)

# 站点统计
station_summary <- df_spatial %>%
  group_by(station, region, lat, lon) %>% 
  summarise(pm25_mean = mean(pm2_5, na.rm = TRUE), .groups = "drop")

# 区域颜色
region_colors <- c("North" = "blue", "South" = "red", "Central" = "orange")
```

## Row {data-height="60%"}

### Interactive Map: Mean PM2.5 by Station

```{r}
pal <- colorNumeric(palette = c("green", "yellow", "red"), domain = station_summary$pm25_mean)
radius_fun <- function(x) scales::rescale(x, to = c(5, 20))

# 区域 Polygon
regions_poly <- df_spatial %>%
  distinct(region, lat, lon) %>%
  group_by(region) %>%
  summarise(poly = list({
    pts <- cbind(lon, lat)
    pts <- pts[complete.cases(pts), ]
    hull <- pts[chull(pts), ]
    hull
  }), .groups = "drop")

map <- leaflet(station_summary) %>% 
  addProviderTiles(providers$CartoDB.Positron)

for (i in 1:nrow(regions_poly)) {
  map <- map %>% addPolygons(
    lng = regions_poly$poly[[i]][,1],
    lat = regions_poly$poly[[i]][,2],
    color = region_colors[regions_poly$region[i]],
    weight = 1,
    fillColor = region_colors[regions_poly$region[i]],
    fillOpacity = 0.15
  )
}

map %>%
  addCircleMarkers(
    lng = ~lon, lat = ~lat,
    radius = ~radius_fun(pm25_mean),
    color = ~pal(pm25_mean),
    fillColor = ~pal(pm25_mean),
    fillOpacity = 0.9,
    stroke = FALSE,
    label = ~paste0("<b>", station, "</b><br>", "Region: ", region, "<br>", "Mean PM2.5: ", round(pm25_mean, 1)),
    labelOptions = labelOptions(direction = "auto", style = list("font-size" = "13px", "background" = "white"))
  ) %>%
  addLegend(pal = pal, values = ~pm25_mean, title = "Mean PM2.5", position = "bottomright")
```

## Row {data-height="40%"}

### PM2.5 Distribution by Region

```{r}
df_spatial %>%
  ggplot(aes(x = region, y = pm2_5, fill = region)) +
  geom_boxplot(outlier.size = 1.2, alpha = 0.75) +
  scale_fill_manual(values = region_colors) +
  theme_minimal(base_size = 13) +
  labs(title = "PM2.5 Distribution Across Regions", y = "PM2.5", x = "Region")
```

### Top 10 Most Polluted Stations

```{r}
DT::datatable(
  station_summary %>% arrange(desc(pm25_mean)) %>% head(10),
  options = list(pageLength = 5, searching = FALSE),
  rownames = FALSE
)
```

### Average PM2.5 by Region

```{r}
region_avg <- station_summary %>%
  group_by(region) %>%
  summarise(avg_pm25 = mean(pm25_mean))

ggplot(region_avg, aes(x = region, y = avg_pm25, fill = region)) +
  geom_col(alpha = 0.8) +
  scale_fill_manual(values = region_colors) +
  theme_minimal(base_size = 13) +
  labs(title = "Average PM2.5 by Region", y = "Average PM2.5", x = "Region")
```

# Meteorological Factors

```{r meteo_prep}
# 准备气象数据
meteo_df <- df_spatial %>%
  dplyr::select(pm2_5, temp, dewp, pres, region, station, lat, lon) %>%
  stats::na.omit()

# 抽样以提高性能
set.seed(123)
n_sample <- min(3000, nrow(meteo_df))
meteo_sample <- meteo_df[sample(seq_len(nrow(meteo_df)), n_sample), ]
meteo_sample$station <- factor(meteo_sample$station)
```

## Row {.tabset}

### PM2.5 vs Temperature

```{r}
plotly::plot_ly(
  data = meteo_sample,
  x = ~temp, y = ~pm2_5,
  type = "scatter", mode = "markers",
  color = ~station,
  text = ~paste("Station:", station, "<br>Region:", region, "<br>Temp:", round(temp, 1), "°C", "<br>PM2.5:", round(pm2_5, 1)),
  hoverinfo = "text",
  marker = list(size = 5, opacity = 0.7)
) %>%
  layout(
    title = "PM2.5 vs Temperature",
    xaxis = list(title = "Temperature (°C)"),
    yaxis = list(title = "PM2.5"),
    legend = list(orientation = "h", x = 0, y = -0.2)
  )
```

### PM2.5 vs Dew Point

```{r}
plotly::plot_ly(
  data = meteo_sample,
  x = ~dewp, y = ~pm2_5,
  type = "scatter", mode = "markers",
  color = ~station,
  text = ~paste("Station:", station, "<br>Region:", region, "<br>Dew Point:", round(dewp, 1), "°C", "<br>PM2.5:", round(pm2_5, 1)),
  hoverinfo = "text",
  marker = list(size = 5, opacity = 0.7)
) %>%
  layout(
    title = "PM2.5 vs Dew Point",
    xaxis = list(title = "Dew Point (°C)"),
    yaxis = list(title = "PM2.5"),
    legend = list(orientation = "h", x = 0, y = -0.2)
  )
```

### PM2.5 vs Pressure

```{r}
plotly::plot_ly(
  data = meteo_sample,
  x = ~pres, y = ~pm2_5,
  type = "scatter", mode = "markers",
  color = ~station,
  text = ~paste("Station:", station, "<br>Region:", region, "<br>Pressure:", round(pres, 1), " hPa", "<br>PM2.5:", round(pm2_5, 1)),
  hoverinfo = "text",
  marker = list(size = 5, opacity = 0.7)
) %>%
  layout(
    title = "PM2.5 vs Pressure",
    xaxis = list(title = "Pressure (hPa)"),
    yaxis = list(title = "PM2.5"),
    legend = list(orientation = "h", x = 0, y = -0.2)
  )
```

# Correlation Analysis

## Row {data-height="35%"}

### Correlation Heatmap

```{r}
cor_mat <- cor(meteo_df[, c("pm2_5", "temp", "dewp", "pres")], use = "complete.obs")
cor_df <- as.data.frame(as.table(cor_mat))
names(cor_df) <- c("Var1", "Var2", "Correlation")

p_cor <- ggplot(cor_df, aes(x = Var1, y = Var2, fill = Correlation)) +
  geom_tile(color = "white") +
  geom_text(aes(label = sprintf("%.2f", Correlation)), size = 3) +
  scale_fill_gradient2(low = "#2166AC", mid = "white", high = "#B2182B", limits = c(-1, 1)) +
  labs(title = "Correlation between PM2.5 and Meteorological Factors", x = NULL, y = NULL, fill = "r") +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1), panel.grid = element_blank())

plotly::ggplotly(p_cor, tooltip = c("x", "y", "fill"))
```

## Row {data-height="65%"}

### Station-wise Correlation Map

```{r}
station_meteo_cor <- df_spatial %>%
  dplyr::group_by(station) %>%
  dplyr::summarise(
    cor_temp = cor(pm2_5, temp, use = "complete.obs"),
    cor_dewp = cor(pm2_5, dewp, use = "complete.obs"),
    cor_pres = cor(pm2_5, pres, use = "complete.obs"),
    n = dplyr::n(),
    .groups = "drop"
  )

station_map <- station_summary %>%
  dplyr::select(station, region, lat, lon) %>%
  dplyr::distinct() %>%
  dplyr::left_join(station_meteo_cor, by = "station")

pal_cor <- colorNumeric(palette = "RdBu", domain = station_map$cor_temp, reverse = TRUE)

leaflet(station_map) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addCircleMarkers(
    lng = ~lon, lat = ~lat,
    radius = 7,
    color = ~pal_cor(cor_temp),
    fillColor = ~pal_cor(cor_temp),
    fillOpacity = 0.9,
    stroke = FALSE,
    popup = ~paste0(
      "<b>", station, "</b><br/>",
      "Corr(PM2.5, Temp) = ", sprintf("%.2f", cor_temp), "<br/>",
      "Corr(PM2.5, Dewp) = ", sprintf("%.2f", cor_dewp), "<br/>",
      "Corr(PM2.5, Pres) = ", sprintf("%.2f", cor_pres), "<br/>",
      "N = ", n
    )
  ) %>%
  addLegend("bottomright", pal = pal_cor, values = ~cor_temp, title = "Corr(PM2.5, Temp)", labFormat = labelFormat(digits = 2))
```

# AQI Analysis

```{r aqi_prep}
# AQI分级
beijing_aqi <- df_spatial %>%
  mutate(
    datetime = as.POSIXct(datetime, format = "%Y-%m-%dT%H:%M:%SZ", tz = "UTC"),
    date = as.Date(datetime)
  ) %>%
  filter(season_year >= 2013, season_year <= 2016)

daily_pm_aqi <- beijing_aqi %>%
  group_by(season_year, date) %>%
  summarise(pm2_5 = mean(pm2_5, na.rm = TRUE), .groups = "drop")

daily_pm_aqi <- daily_pm_aqi %>%
  mutate(
    aqi_level = case_when(
      pm2_5 <= 35  ~ "Level 1 (Good)",
      pm2_5 <= 75  ~ "Level 2 (Moderate)",
      pm2_5 <= 115 ~ "Level 3 (Unhealthy-Sensitive)",
      pm2_5 <= 150 ~ "Level 4 (Unhealthy)",
      pm2_5 <= 250 ~ "Level 5 (Very Unhealthy)",
      pm2_5 > 250  ~ "Level 6 (Hazardous)",
      TRUE ~ NA_character_
    ),
    aqi_level = factor(aqi_level, levels = c("Level 1 (Good)", "Level 2 (Moderate)", "Level 3 (Unhealthy-Sensitive)", "Level 4 (Unhealthy)", "Level 5 (Very Unhealthy)", "Level 6 (Hazardous)"))
  )

aqi_prop <- daily_pm_aqi %>%
  filter(!is.na(aqi_level)) %>%
  count(season_year, aqi_level) %>%
  group_by(season_year) %>%
  mutate(prop = n / sum(n)) %>%
  ungroup()

aqi_colors <- c(
  "Level 1 (Good)" = "#2ECC71",
  "Level 2 (Moderate)" = "#F1C40F",
  "Level 3 (Unhealthy-Sensitive)" = "#E67E22",
  "Level 4 (Unhealthy)" = "#E74C3C",
  "Level 5 (Very Unhealthy)" = "#8E44AD",
  "Level 6 (Hazardous)" = "#7F0000"
)

waffle_df <- aqi_prop %>%
  mutate(n_cells = round(prop * 100)) %>%
  filter(n_cells > 0) %>%
  tidyr::uncount(n_cells) %>%
  group_by(season_year) %>%
  mutate(
    cell_index = row_number() - 1,
    x = cell_index %% 10,
    y = cell_index %/% 10
  ) %>%
  ungroup()
```

## Row {data-height="35%"}

### AQI Severity Distribution by Seasonal Year

```{r}
p_stack <- ggplot(aqi_prop, aes(x = factor(season_year), y = prop, fill = aqi_level)) +
  geom_col(width = 0.7, colour = "white") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  scale_fill_manual(values = aqi_colors, name = "AQI level") +
  labs(
    title = "AQI Severity Distribution by Seasonal Year",
    subtitle = "Daily mean PM2.5 aggregated over all stations (Mar–Feb)",
    x = "Seasonal year",
    y = "Proportion of days"
  ) +
  theme_minimal() +
  theme(legend.position = "right")

plotly::ggplotly(p_stack)
```

## Row {data-height="65%"}

### AQI Severity Waffle Chart

```{r}
ggplot(waffle_df, aes(x = x, y = y, fill = aqi_level)) +
  geom_tile(colour = "white", linewidth = 0.5) +
  scale_y_reverse() +
  coord_equal() +
  scale_fill_manual(values = aqi_colors, name = "AQI level") +
  facet_wrap(~ season_year, nrow = 1) +
  labs(
    title = "AQI Severity Waffle Chart by Seasonal Year",
    subtitle = "Each grid ≈ 1% of days in the seasonal year"
  ) +
  theme_void() +
  theme(
    strip.text = element_text(size = 12, face = "bold"),
    legend.position = "right",
    legend.text = element_text(size = 10),
    plot.title = element_text(size = 14, face = "bold"),
    plot.subtitle = element_text(size = 11)
  )
```

# Wind Patterns

```{r wind_prep}
wd_levels_full <- c("N","NNE","NE","ENE","E","ESE","SE","SSE","S","SSW","SW","WSW","W","WNW","NW","NNW","CV")
aqi_palette <- c("darkgreen", "gold", "orange", "red", "purple", "brown")

wd_full_names <- c(
  N = "North", NNE = "North-Northeast", NE = "Northeast", ENE = "East-Northeast",
  E = "East", ESE = "East-Southeast", SE = "Southeast", SSE = "South-Southeast",
  S = "South", SSW = "South-Southwest", SW = "Southwest", WSW = "West-Southwest",
  W = "West", WNW = "West-Northwest", NW = "Northwest", NNW = "North-Northwest"
)

# 准备风向数据
wind_data <- df_spatial %>%
  filter(!is.na(wd), !is.na(aqi_level)) %>%
  mutate(wd = factor(wd, levels = wd_levels_full))
```

## Row {data-height="50%"}

### Pollution Rose (All Periods)

```{r}
df_rose_all <- wind_data %>%
  filter(wd != "CV") %>%
  count(wd, aqi_level, name = "n") %>%
  group_by(wd) %>%
  mutate(freq = n / sum(n)) %>%
  ungroup()

if (nrow(df_rose_all) > 0) {
  ggplot(df_rose_all, aes(x = wd, y = freq, fill = aqi_level)) +
    geom_col(width = 1, color = "gray25") +
    coord_polar(start = -pi/16) +
    scale_y_continuous(labels = percent_format(accuracy = 1)) +
    scale_fill_manual(values = aqi_palette, name = "AQI level") +
    labs(title = "Pollution Rose (All Periods)", subtitle = "Frequency share within direction", x = "Wind direction", y = "Frequency") +
    theme_minimal() +
    theme(legend.position = "bottom")
}
```

### Winter vs Summer Comparison

```{r}
df_rose_ws <- wind_data %>%
  filter(season %in% c("Winter", "Summer"), wd != "CV") %>%
  count(season, wd, aqi_level, name = "n") %>%
  group_by(season, wd) %>%
  mutate(freq = n / sum(n)) %>%
  ungroup()

if (nrow(df_rose_ws) > 0) {
  ggplot(df_rose_ws, aes(x = wd, y = freq, fill = aqi_level)) +
    geom_col(width = 1, color = "gray25") +
    coord_polar(start = -pi/16) +
    scale_y_continuous(labels = percent_format(accuracy = 1)) +
    scale_fill_manual(values = aqi_palette, name = "AQI level") +
    labs(title = "Winter vs Summer: Pollution Rose", subtitle = "Frequency share within direction", x = "Wind direction", y = "Frequency") +
    facet_wrap(~season, ncol = 2) +
    theme_minimal() +
    theme(legend.position = "bottom")
}
```

## Row {data-height="50%"}

### Mean PM2.5 by Wind Direction (colored by wind speed)

```{r}
df_pm_dir <- wind_data %>%
  filter(wd != "CV", !is.na(pm2_5), !is.na(wspm)) %>%
  group_by(wd) %>%
  summarize(
    mean_pm25 = mean(pm2_5, na.rm = TRUE),
    mean_wspm = mean(wspm, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(dir_full = factor(recode(wd, !!!wd_full_names), levels = wd_full_names[wd_levels_full[wd_levels_full != "CV"]]))

if (nrow(df_pm_dir) > 0) {
  p_dir <- ggplot(df_pm_dir, aes(
    x = dir_full,
    y = mean_pm25,
    fill = mean_wspm,
    text = paste0("Direction: ", dir_full, "<br>", "Mean PM2.5: ", round(mean_pm25, 1), " µg/m³<br>", "Mean wind speed: ", round(mean_wspm, 2), " m/s")
  )) +
    geom_col(color = "gray25") +
    scale_fill_viridis_c(name = "Wind speed (avg)") +
    labs(title = "Mean PM2.5 by Wind Direction", subtitle = "Fill = mean wind speed (m/s)", x = "Wind direction", y = "Mean PM2.5 (µg/m³)") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "right")
  
  ggplotly(p_dir, tooltip = "text")
}
```

# Data Tables

## Row

### Raw Data Sample

```{r}
DT::datatable(
  aq %>% select(datetime, station, region, pm2_5, temp, dewp, pres, wd, wspm, season) %>% head(100),
  options = list(pageLength = 10, scrollX = TRUE),
  filter = "top",
  rownames = FALSE
)
```

# Reference

## Column

### About flexdashboard

This dashboard demonstrates the capabilities of **flexdashboard** for R:

-   **Layout options**: rows, columns, pages, tabs, and sidebars
-   **Interactive visualizations**: highcharter, plotly, leaflet
-   **Responsive design**: adapts to different screen sizes
-   **Easy publishing**: can be hosted on any web server

### Resources

-   [flexdashboard documentation](https://pkgs.rstudio.com/flexdashboard/)
-   [Icon sets](https://pkgs.rstudio.com/flexdashboard/articles/using.html#icon-sets)
-   [Themes](https://pkgs.rstudio.com/flexdashboard/articles/theme.html)

### Data Source

**Beijing Multi-Site Air-Quality Data**

-   Period: 2013-2017
-   Monitoring stations: 12 sites across Beijing
-   Parameters: PM2.5, PM10, SO2, NO2, CO, O3, temperature, pressure, wind, etc.
