---
title: "R Tutorial - Exercise"
author: "Junya YANG 2230026188"
date: "2025 Sep 20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Q1. Let $x$ be defined as follow.\
a. Add 3 to just the even-index elements.\
b. Compute the square root of each element.

```{r}
x <- c(11,10,5,9,14,12,18,2,17,9,1,7,1,15,16,7,16,12,13,3,8,3,11,1,5,7,7,10,6,10,17,13,1,2,17,1,15,8,12,3,6,15,10,3,10,12,19,17,15,12,18,17,1,13,5,9,18,2,20,16,15,7,9,12,12,5,5,7,7,9,17,7,15,11,10,12,19,8,13,3,7,14,13,18,1,11,18,1,12,18,5,12,5,10,3,6,11,12,1,3)
```

```{r}
# Write your answer here
#len = length(x)
#even_idx = seq(from=2,to =len,by=2)
#x[even_idx] = x[even_idx]+3
x[c(FALSE, TRUE)]=x[c(FALSE, TRUE)]+3
print(x)
x=sqrt(x)
print(x)
```

Q2. Given x = [3 15 9 12 -1 -12 9 6 1], provide the command that will\
a) set the values of x that are positive to zero\
b) set values that are multiples of 3 to 3\
c) multiply the values of x that are even by 5\
d) extract the values of x that are greater than 10 into a vector called y\
e) Find the index position of elements in x which are larger than 4

```{r}
# Write your answer here
# a) set the values of x that are positive to zero
x =c(3,15,9,12,-1,-12,9,6,1)
x[x>0]=0
print(x)
# b) set values that are multiples of 3 to 3
x =c(3,15,9,12,-1,-12,9,6,1)
x[x%%3==0]=3
print(x)
# c) multiply the values of x that are even by 5
x =c(3,15,9,12,-1,-12,9,6,1)
x[x%%2==0]=x[x%%2==0]*5
print(x)
# d) extract the values of x that are greater than 10 into a vector called y
x =c(3,15,9,12,-1,-12,9,6,1)
y=x[x>10]
print(y)
# e) Find the index position of elements in x which are larger than 4
x =c(3,15,9,12,-1,-12,9,6,1)
len = length(x)
list_idx = c(1:len)
idx = list_idx[x>4]
#idx <- which(x > 4)
print(idx)

```

Q3. Given the matrix A as following

```{r}
a <- c(2,4,1,6,7,2,3,5,9)
A <- matrix(a, nrow=3, ncol=3, byrow=T)
A
```

provide the commands needed to\
a) assign the last 2 rows of A to a matrix called y\
b) compute the sum over the columns of A\
c) compute the sum over the rows of A\
d) compute the standard deviation of each column of A\
e) Standardize the matrix A such that columns of A are centered to have mean 0 and scaled to have standard deviation 1.

```{r}
# Write your answer here
# a) assign the last 2 rows of A to a matrix called y
y = tail(A,2)
print(y)
# b) compute the sum over the columns of A
col_sum = apply(A,2,sum) # Option 1
print(colSums(A)) # Option 2
# c) compute the sum over the rows of A
row_sum = apply(A,1,sum) 
print(rowSums(A))
# d) compute the standard deviation of each column of A
col_sd = apply(A, 2, sd)
print(col_sd)
# e) Standardize the matrix A such that columns of A are centered to have mean 0 and scaled to have standard deviation 1.
print(scale(A))

```

Q4. The data file “country.txt” in Ispace contains numerous population indicators for a sample of 115 countries. Read the data file into R and answer the following questions.\
a. How many variables are included in the dataset?\
b. What is the percentage of developing country (develop=1) in this data?\
c. What is the standard deviation of the variable death rate?\
d. What is the range of GDP for the countries in which death rate is greater than 8?\
e. What is the mean of GDP for the developing countries which have death rate higher than 10?\
f. What is the correlation coefficient between GDP and birthrate? Discuss the relationship between GDP and the birthrate.

```{r}
# Write your answer here
# a. How many variables are included in the dataset?
data <- read.table(file = "./country.txt",header=T)
#print(ncol(data))
x=dim(data)
print(x[2])
# b. What is the percentage of developing country (develop=1) in this data?
num_developing=sum(data$develop==1)
total_countries = nrow(data)
percent = (num_developing/total_countries)*100
print(percent)
# c. What is the standard deviation of the variable death rate?
sd_death = sd(data$deathrate)
print(sd_death)
# d. What is the range of GDP for the countries in which death rate is greater than 8?
#countries=data[data$deathrate>8,"gdp"]
countries=data[data$deathrate>8,]
print(range(countries$gdp))
#print(min(countries$gdp))
#print(max(countries$gdp))
# e. What is the mean of GDP for the developing countries which have death rate higher than 10?
print(mean(data[data$develop==1 & data$deathrate>10,"gdp"]))
# f. What is the correlation coefficient between GDP and birthrate? Discuss the relationship between GDP and the birthrate.
print(cor(data$gdp, data$birthrate))
print("It shows that there is a negative correlation between GDP and birthrate. This means that, in general, the higher a country's GDP, the lower its birth rate.")

```

Q5. Fibonacci sequence is defined by F(1)=F(2)=1 and F(n)=F(n–1)+F(n–2). Write an R function called `Fibonacci` which displays the first n Fibonacci numbers and computes their average. For example, if the input n is 20, the output of Fibonacci (20) shall look like:\
The first 20 Fibonacci numbers are:\
1 1 2 3 5 8 13 21 34 55 89 144 233 377 610 987 1597 2584 4181 6765\
The average is 885.5

Fill the blank in the code, then delete "eval=FALSE, message=TRUE" to run.

```{r}
Fibonacci <- function(n) {
  if (n == 1) {
    F <- c(1)
  } else {
    F <- c(1, 1)
  }
  if (n > 2) {
    for (i in 3:n) {
      next_num <- F[i-1] + F[i-2]
      F <- c(F, next_num)
    }
  }
  af <- mean(F)
  cat(paste('The first', n, 'Fibonacci numbers are:'), '\n', F, '\n', paste('the average is', af))
}

Fibonacci(20)
```
